---
title: "Lab 1: Introduction to Field Spectroscopy"
subtitle: "Analyzing Spectral Reflectance"
author: "Ronny A. Hernández Mora"
date: "02/10/2021"
output:
  rmarkdown::pdf_document:
    fig_caption: yes        
    includes:  
      in_header: my_header.tex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(dplyr)
library(ggplot2)
library(readr)
library(fs)
library(purrr)
library(stringr)
library(gt)
```

<!-- Import leaf reflectance data -->
```{r leaf_reflectance_data}
files <- dir_ls(path = "data", glob = "data/reflectance*")

reflectance <- map_dfr(.x = files, 
                       .id = "file_id",
                       .f = function(archivo) {
                         read_csv(archivo,
                                  skip = 4,
                                  col_names = FALSE) %>% 
                           rename(wavelength = X1,
                                  reflectance = X2)
                         
                       }) %>% 
  mutate(file_id = str_extract(file_id, "(-[^.]+)"))
```

<!-- Importa leaf first derivative data -->
```{r leaf_first_derivative_data}
files <- dir_ls(path = "data",  glob = "data/first_derivative*")

first_derivative <- map_dfr(.x = files, 
                            .id = "file_id",
                            .f = function(archivo) {
                              read_csv(archivo,
                                       skip = 4,
                                       col_names = FALSE) %>% 
                                rename(wavelength = X1,
                                       first_derivative = X2)
                              
                            }) %>% 
  mutate(file_id = str_extract(file_id, "(_[^.]+)"))
```

<!-- Importa leaf ndvi data -->
```{r leaf_ndvi_data}
files <- dir_ls(path = "data", glob = "data/ndvi*")

leaves_ndvi <- map_dfr(.x = files, 
                       .id = "file_id",
                       .f = function(archivo) {
                         read_csv(archivo)   
                       }) %>% 
  mutate(file_id = str_extract(file_id, "(_[^.]+)"))
```

<!-- Import random sample data -->
```{r random_sample_data}
## Random samples ---
ratio_06 <- read_csv("data/random_06_ratios.csv")
ratio_18 <- read_csv("data/random_18_ratios.csv")
ratio_23 <- read_csv("data/random_23_ratios.csv")

random_ratios <- bind_rows(ratio_06, ratio_18, ratio_23)
```

<!-- Import random first derivative -->

```{r}
first_d_06 <- read_csv("data/random_06_first_derivative.csv",
                       skip = 4,
                       col_names = FALSE) %>% 
  mutate(id = "random_06")

first_d_18 <- read_csv("data/random_18_first_derivative.csv",
                       skip = 4,
                       col_names = FALSE) %>% 
  mutate(id = "random_18")


first_d_23 <- read_csv("data/random_23_first_derivative.csv",
                       skip = 4,
                       col_names = FALSE) %>% 
  mutate(id = "random_23")

random_first_derivative <- bind_rows(first_d_06, 
                                     first_d_18,
                                     first_d_23) %>% 
  rename("wavelength" = "X1",
         "first_derivative" = "X2")
```

<!-- Import random second derivative -->

```{r}
second_d_06 <- read_csv("data/random_06_second_derivative.csv",
                       skip = 4,
                       col_names = FALSE) %>% 
  mutate(id = "random_06")

second_d_18 <- read_csv("data/random_18_second_derivative.csv",
                       skip = 4,
                       col_names = FALSE) %>% 
  mutate(id = "random_18")

second_d_23 <- read_csv("data/random_23_second_derivative.csv",
                       skip = 4,
                       col_names = FALSE) %>% 
  mutate(id = "random_23")

random_second_derivative <- bind_rows(second_d_06, 
                                      second_d_18,
                                      second_d_23) %>% 
  rename("wavelength" = "X1",
         "second_derivative" = "X2")
```

## Introduction

<!-- -   Theoretical Background -- Define the problem and how you attempt to solve it. -->

<!-- -   Define the terminology you will be using. -->

<!--       1. Spectral Signature: -->
<!--       2. Red Edge: -->
<!--       3. Absorption Feature: -->
<!--       4. Directional and Diffuse Light: -->
<!--       5. First Derivative: -->

<!-- -   State the objective and purpose of the lab. -->

## Methods

 <!-- - Discuss any techniques/tools used in the lab. Why are these -->
 <!-- relevant? -->
 <!-- - What processes are involved? Illustrate the methods via a -->
 <!-- flowchart. -->

## Results

 <!-- - Quantitatively and qualitatively present your results. -->
 
### Random samples

#### Random samples ratios

<!-- select 3 ratios -->
<!-- PRI = Photochemical reflectance index (visible ratio)-->
<!-- https://www.sciencedirect.com/science/article/abs/pii/S0034425720302583 -->

<!-- NDVI = normalized difference vegetation index (visible/NIR ratio) -->

```{r random_ratios, fig.width = 9}
random_ratios %>%
  select(`File name`, `PRI (531-570)/(531+570)`, 
         `NDVI (774-677)/(774+677)`, `NDVI (800-680)/(800+680)`) %>%
  gt() %>% 
  tab_header(title = "Selected ratios values for the random samples") %>% 
  tab_source_note(
    source_note = "Table 1. Values for the selected ratios obtained from the
                  random samples with the spectrometer UniSpec SC"
  )
```

#### Random samples wavelengths

<!-- fig.cap = 'First line of a figure caption, \\\\ followed by an incredibly, -->

#### Random samples derivatives

```{r random_first_derivative, fig.cap = "Values for the random spectra (first derivative) of three samples. The selected samples file names choosen are 001-00006.spu (random_06), 001-00018.spu (random_18) and 001-00023 (random_23)  "}
random_first_derivative %>% 
  filter(wavelength > 400, wavelength < 900) %>%
  ggplot(aes(x = wavelength, 
             y = first_derivative, 
             color = id)) +
  geom_line() +
  scale_y_continuous(breaks = seq(-0.009, 0.025, by = 0.002)) +
  scale_x_continuous(breaks = seq(400, 900, by = 50)) +
  scale_color_viridis_d() +
  labs(x = "Wavelength (nm)",
       y = "First derivative",
       color = "Random sample") +
  theme_light(base_size = 11)
```


```{r random_second_derivative, fig.cap = "Values for the random spectra (second derivative) of three samples. The selected samples file names choosen are 001-00006.spu (random_06), 001-00018.spu (random_18) and 001-00023 (random_23)  "}
random_second_derivative %>% 
  filter(wavelength > 400, wavelength < 900) %>%
  ggplot(aes(x = wavelength, 
             y = second_derivative, 
             color = id)) +
  geom_line() +
  scale_y_continuous(breaks = seq(-0.008, 0.008, by = 0.001)) +
  scale_x_continuous(breaks =  seq(400, 900, by = 50)) +
  scale_color_viridis_d() +
  labs(x = "Wavelength",
       y = "Second derivative",
       color = "Random sample") +
  theme_light(base_size = 11)
```

### Leaves reflectances

```{r leaf_reflectances, fig.cap = "Reflectance for four leaf samples with different senescence state. Green and yellow 1 leaves shows more absorbance in the 650 nm to 700 nm compare too the red and yellow 2 samples."}
reflectance %>% 
  filter(wavelength > 400, wavelength < 900) %>% 
  ggplot(aes(x = wavelength, y = reflectance, color = file_id)) +
  geom_line() +
  scale_color_viridis_d() +
  scale_y_continuous(breaks = seq(0, 1.5, by = 0.1)) +
  scale_x_continuous(breaks = seq(400, 1000, by = 50)) +
  labs(x = "Wavelength (nm)",
       y = "Reflectance",
       color = "Leaf id") +
  theme_light(base_size = 11)
```

### Leaves first derivative

```{r leaf derivative, fig.cap = "Spectra (first derivative) for four leaf samples with different senescence state."}
first_derivative %>% 
  filter(wavelength > 400, wavelength < 1000) %>% 
  ggplot(aes(x = wavelength, y = first_derivative, color = file_id)) +
  geom_line() +
  scale_y_continuous(breaks = seq(-0.009, 0.03, by = 0.002)) +
  scale_x_continuous(breaks = seq(400, 900, by = 50)) +
  scale_color_viridis_d(labels = c("Green leaf", "Red leaf",
                                   "Yellow 1 leaf", "Yellow 2 leaf")) +
  labs(x = "Wavelength (nm)",
       y = "First derivative",
       color = "Leaf id") +
  theme_light(base_size = 11)
```

### Leaves NDVIs

```{r leaves_ndvis}
leaves_ndvi %>%
  select(`File name`, `NDVI (774-677)/(774+677)`) %>%
  gt() %>% 
  tab_header(title = "NDVI values for green, red and yellows leaves") %>% 
  tab_source_note(
    source_note = "Table 2. Normalized difference vegetation index values (667-774 nm) for the sampled leaves"
  )
```

## Discussion / conclussion

 <!-- - What do the results mean? -->
 <!-- - Was the original problem resolved effectively? -->
 <!-- - What can this information be used for? -->


## Lab questions

  1. Explain some of the errors and considerations involved with the collection 
  of spectra. (eg. Light Conditions (diffuse, directional)). Provide 3 examples
  and briefly explain how they influence the data collection.
  2. List the main biochemical drivers of leaf spectra. (3 minimum) (e.g:
  chlorophyl A)
  3. If the reflectance is a value of percentage (%) based on the incoming 
  light, what does a value greater than 1 (100%) imply? (Think about the signal)
