---
title: "Lab 1: Introduction to Field Spectroscopy"
subtitle: "Analyzing Spectral Reflectance"
author: "Ronny A. Hernández Mora"
date: "02/10/2021"
output:
  rmarkdown::pdf_document:
    fig_caption: yes        
    includes:  
      in_header: my_header.tex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(readxl)
library(janitor)
library(dplyr)
library(ggplot2)
library(readr)
library(fs)
library(purrr)
library(stringr)
library(gt)
```

<!-- Import leaf reflectance data -->
```{r leaf_reflectance_data}
files <- dir_ls(path = "data", glob = "data/reflectance*")

reflectance <- map_dfr(.x = files, 
                       .id = "file_id",
                       .f = function(archivo) {
                         read_csv(archivo,
                                  skip = 4,
                                  col_names = FALSE) %>% 
                           rename(wavelength = X1,
                                  reflectance = X2)
                         
                       }) %>% 
  mutate(file_id = str_extract(file_id, "(-[^.]+)"))
```

<!-- Importa leaf first derivative data -->
```{r leaf_first_derivative_data}
files <- dir_ls(path = "data",  glob = "data/first_derivative*")

first_derivative <- map_dfr(.x = files, 
                            .id = "file_id",
                            .f = function(archivo) {
                              read_csv(archivo,
                                       skip = 4,
                                       col_names = FALSE) %>% 
                                rename(wavelength = X1,
                                       first_derivative = X2)
                              
                            }) %>% 
  mutate(file_id = str_extract(file_id, "(_[^.]+)"))
```

<!-- Importa leaf ndvi data -->
```{r leaf_ndvi_data}
files <- dir_ls(path = "data", glob = "data/ndvi*")

leaves_ndvi <- map_dfr(.x = files, 
                       .id = "file_id",
                       .f = function(archivo) {
                         read_csv(archivo)   
                       }) %>% 
  mutate(file_id = str_extract(file_id, "(_[^.]+)"))
```

<!-- Import random sample data -->
```{r random_sample_data}
## Random samples ---
ratio_06 <- read_csv("data/random_06_ratios.csv")
ratio_18 <- read_csv("data/random_18_ratios.csv")
ratio_23 <- read_csv("data/random_23_ratios.csv")

random_ratios <- bind_rows(ratio_06, ratio_18, ratio_23)
```

<!-- Import random first derivative -->

```{r}
first_d_06 <- read_csv("data/random_06_first_derivative.csv",
                       skip = 4,
                       col_names = FALSE) %>% 
  mutate(id = "random_06")

first_d_18 <- read_csv("data/random_18_first_derivative.csv",
                       skip = 4,
                       col_names = FALSE) %>% 
  mutate(id = "random_18")


first_d_23 <- read_csv("data/random_23_first_derivative.csv",
                       skip = 4,
                       col_names = FALSE) %>% 
  mutate(id = "random_23")

random_first_derivative <- bind_rows(first_d_06, 
                                     first_d_18,
                                     first_d_23) %>% 
  rename("wavelength" = "X1",
         "first_derivative" = "X2")
```

<!-- Import random second derivative -->

```{r}
second_d_06 <- read_csv("data/random_06_second_derivative.csv",
                        skip = 4,
                        col_names = FALSE) %>% 
  mutate(id = "random_06")

second_d_18 <- read_csv("data/random_18_second_derivative.csv",
                        skip = 4,
                        col_names = FALSE) %>% 
  mutate(id = "random_18")

second_d_23 <- read_csv("data/random_23_second_derivative.csv",
                        skip = 4,
                        col_names = FALSE) %>% 
  mutate(id = "random_23")

random_second_derivative <- bind_rows(second_d_06, 
                                      second_d_18,
                                      second_d_23) %>% 
  rename("wavelength" = "X1",
         "second_derivative" = "X2")
```


<!-- Import Rocks data -->

```{r}
files <- dir_ls(path = "data/", glob = "data/rock*")

rock_1 <- map_dfr(.x = files, 
                  .id = "file_id",
                  .f = function(archivo) {
                    read_csv(archivo,
                             skip = 4,
                             col_names = FALSE) %>% 
                      rename(wavelength = X1,
                             reflectance = X2)
                    
                  }) 
```



# Introduction

<!-- -   Theoretical Background -- Define the problem and how you attempt to solve it. -->

<!-- -   Define the terminology you will be using. -->

<!--       1. Spectral Signature: -->
<!--       2. Red Edge: -->
<!--       3. Absorption Feature: -->
<!--       4. Directional and Diffuse Light: -->
<!--       5. First Derivative: -->

<!-- -   State the objective and purpose of the lab. -->

# Methods

 <!-- - Discuss any techniques/tools used in the lab. Why are these -->
 <!-- relevant? -->
 <!-- - What processes are involved? Illustrate the methods via a -->
 <!-- flowchart. -->

# Results

 <!-- - Quantitatively and qualitatively present your results. -->
 
## Random samples

In the case of random samples, we present the results for ratios, wavelengths 
and the first and second derivative. Each section result is showed below.

### Random samples ratios

<!-- select 3 ratios -->
<!-- PRI = Photochemical reflectance index (visible ratio)-->
<!-- https://www.sciencedirect.com/science/article/abs/pii/S0034425720302583 -->

<!-- NDVI = normalized difference vegetation index (visible/NIR ratio) -->

```{r random_ratios}
random_ratios %>%
  select(`File name`, `PRI (531-570)/(531+570)`, 
         `NDVI (774-677)/(774+677)`, `NDVI (800-680)/(800+680)`) %>%
  gt() %>% 
  tab_options(
    table.font.size = 12,
    table.width = 10
  ) %>% 
  # tab_style(
  #   style = cell_text(size = px(12)),
  #   locations = 
  #   ) %>% 
  tab_header(title = "Selected ratios values for the random samples") %>% 
  tab_source_note(
    source_note = "Table 1. Values for the selected ratios obtained from the
                  random samples with the spectrometer UniSpec SC"
  )
```

### Random samples wavelengths

<!-- fig.cap = 'First line of a figure caption, \\\\ followed by an incredibly, -->

### Random samples derivatives

```{r random_first_derivative, fig.cap = "Values for the random spectra (first derivative) of three samples. The selected samples file names choosen are 001-00006.spu (random_06), 001-00018.spu (random_18) and 001-00023 (random_23)  "}
random_first_derivative %>% 
  filter(wavelength > 400, wavelength < 900) %>%
  ggplot(aes(x = wavelength, 
             y = first_derivative, 
             color = id)) +
  geom_line() +
  scale_y_continuous(breaks = seq(-0.009, 0.025, by = 0.002)) +
  scale_x_continuous(breaks = seq(400, 900, by = 50)) +
  scale_color_viridis_d() +
  labs(x = "Wavelength (nm)",
       y = "First derivative",
       color = "Random sample") +
  theme_light(base_size = 11)
```


```{r random_second_derivative, fig.cap = "Values for the random spectra (second derivative) of three samples. The selected samples file names choosen are 001-00006.spu (random_06), 001-00018.spu (random_18) and 001-00023 (random_23)  "}
random_second_derivative %>% 
  filter(wavelength > 400, wavelength < 900) %>%
  ggplot(aes(x = wavelength, 
             y = second_derivative, 
             color = id)) +
  geom_line() +
  scale_y_continuous(breaks = seq(-0.008, 0.008, by = 0.001)) +
  scale_x_continuous(breaks =  seq(400, 900, by = 50)) +
  scale_color_viridis_d() +
  labs(x = "Wavelength",
       y = "Second derivative",
       color = "Random sample") +
  theme_light(base_size = 11)
```

## Leaves

### Leaves reflectances

```{r leaf_reflectances, fig.cap = "Reflectance for four leaf samples with different senescence state. Green and yellow 1 leaves shows more absorbance in the 650 nm to 700 nm compare too the red and yellow 2 samples."}
reflectance %>% 
  filter(wavelength > 400, wavelength < 900) %>% 
  ggplot(aes(x = wavelength, y = reflectance, color = file_id)) +
  geom_line() +
  scale_color_viridis_d() +
  scale_y_continuous(breaks = seq(0, 1.5, by = 0.1)) +
  scale_x_continuous(breaks = seq(400, 1000, by = 50)) +
  labs(x = "Wavelength (nm)",
       y = "Reflectance",
       color = "Leaf id") +
  theme_light(base_size = 11)
```

### Leaves first derivative

```{r leaf derivative, fig.cap = "Spectra (first derivative) for four leaf samples with different senescence state."}
first_derivative %>% 
  filter(wavelength > 400, wavelength < 1000) %>% 
  ggplot(aes(x = wavelength, y = first_derivative, color = file_id)) +
  geom_line() +
  scale_y_continuous(breaks = seq(-0.009, 0.03, by = 0.002)) +
  scale_x_continuous(breaks = seq(400, 900, by = 50)) +
  scale_color_viridis_d(labels = c("Green leaf", "Red leaf",
                                   "Yellow 1 leaf", "Yellow 2 leaf")) +
  labs(x = "Wavelength (nm)",
       y = "First derivative",
       color = "Leaf id") +
  theme_light(base_size = 11)
```

### Leaves NDVIs

```{r leaves_ndvis}
leaves_ndvi %>%
  select(`File name`, `NDVI (774-677)/(774+677)`) %>%
  gt() %>% 
  tab_header(title = "NDVI values for green, red and yellows leaves") %>% 
  tab_source_note(
    source_note = "Table 2. Normalized difference vegetation index values (667-774 nm) for the sampled leaves"
  )
```

## Rocks

### Rocks reflectances

```{r rock_reflectance, fig.cap = "Reflectance of rock 1."}
rock_1 %>% 
  filter(wavelength > 400, wavelength < 900) %>% 
  ggplot(aes(x = wavelength, y = reflectance, color = file_id)) +
  geom_line() +
  scale_color_viridis_d() +
  scale_y_continuous(breaks = seq(0, 1.5, by = 0.1)) +
  scale_x_continuous(breaks = seq(400, 1000, by = 50)) +
  labs(x = "Wavelength (nm)",
       y = "Reflectance",
       color = "Leaf id") +
  theme_light(base_size = 11)
```


# Discussion / conclussion

 <!-- - What do the results mean? -->
 <!-- - Was the original problem resolved effectively? -->
 <!-- - What can this information be used for? -->


# Lab questions

  1. Explain some of the errors and considerations involved with the collection 
  of spectra. (eg. Light Conditions (diffuse, directional)). Provide 3 examples
  and briefly explain how they influence the data collection.
  2. List the main biochemical drivers of leaf spectra. (3 minimum) (e.g:
  chlorophyl A)
  3. If the reflectance is a value of percentage (%) based on the incoming 
  light, what does a value greater than 1 (100%) imply? (Think about the signal)
  4. Provide a graph which includes the ‘% reflectance, ‘reference’ and ‘dark
  current’. The graph must have appropriate axis values (400,1100 & 0,100 for 
  X-Y axis respectively), the graph must also have titles, labels and display 
  the legend. ***
  5. Estimate an approximate NDVI (Normalized Difference Vegetation Index) for 
  this Unknown Spectra. (Red can be an average between 625nm to 700nm, NIR can 
  be an average between 750nm to 900nm.) Recall that the formula is: 
  NDVI = (NIR – RED)/(NIR + RED)





```{r}
# Function to calculate reflectance as indicated in lab guide.
calculate_reflectance <- function(sample_value, dark_value, reference_value) {
  
  sample_dark <- sample_value - dark_value
  reference_dark <- reference_value - dark_value
  
  reflectance <- (sample_dark / reference_dark) * 100
  
  return(reflectance)
  
}

# Import data and mutate variables
spectra_correction <- read_xlsx("data/Spectra_correction_example.xlsx") %>% 
  clean_names() %>% 
  mutate(average = (measure_1 + measure_2 + measure_3 + measure_4) / 4,
         reflectance = calculate_reflectance(sample_value = average, 
                                             dark_value = dark_current,
                                             reference_value = reference))
```


```{r}
ggplot(data = spectra_correction) +
  geom_line(aes(x = wavelength_nm, y = average, color = "Reflectance")) +
  geom_line(aes(x = wavelength_nm, y = reference, color = "Reference")) +
  geom_line(aes(x = wavelength_nm, y = dark_current, color = "Dark current")) +
  scale_color_viridis_d() +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  scale_x_continuous(breaks = seq(400, 1000, by = 50)) +
  labs(x = "Wavelength (nm)",
       y = "Reflectance",
       color = "Type of measurement") +
  theme_light(base_size = 11)




```

